Uncertainty propagation for predictor effects and bias correction in generalized linear (mixed) models

### 2021-12-13 variable_predictions.Rnw
Autocommit

@@

We consider more complex interactions in which the focal predictor is modelled as a cubic polynomial. Specifically:
%
\begin{align}\label{sim:lm_cubic}
\mathrm{hh~size} &= \beta_0 + \beta_{\mathrm{A_1}}\mathrm{Age}_i + \beta_{\mathrm{A_2}}\mathrm{Age}^2_i + \beta_{\mathrm{A_3}}\mathrm{Age}^3_i + \beta_{\mathrm{W}}\mathrm{Wealthindex}_i + \epsilon_i \nonumber\\
\mathrm{Age}_i &\sim \mathrm{Normal}(0, 1) \nonumber\\
\mathrm{Wealthindex}_i &\sim \mathrm{Normal}(0, 1) \nonumber\\
\epsilon_i &\sim \mathrm{Normal}(0, 10) \nonumber\\
\beta_0 &= 20 \nonumber\\
\beta_{\mathrm{A}_1} &= 0.1 \nonumber\\
\beta_{\mathrm{A}_2} &= 0.8 \nonumber\\
\beta_{\mathrm{A}_3} &= 0.3 \nonumber\\
\beta_{\mathrm{W}} &= 0.8 \nonumber\\
i &= 1,\cdots, 100
\end{align}
%

>>>>

\newcommand{\yE}{{\mathrm E}}
\let\over=\overline

@@

We consider more complex interactions in which the focal predictor is modelled as a cubic polynomial. Specifically:
%
\begin{align}\label{sim:lm_cubic}
\mathrm{hh~size} &= \beta_0 + \beta_{\mathrm{A_1}}\mathrm{Age}_i + \beta_{\mathrm{A_2}}\mathrm{Age}^2_i + \beta_{\mathrm{A_3}}\mathrm{Age}^3_i + \beta_{\mathrm{W}}\mathrm{Wealthindex}_i + \epsilon_i \nonumber\\
\mathrm{Age}_i &\sim \mathrm{Normal}(0, 1) \nonumber\\
\mathrm{Wealthindex}_i &\sim \mathrm{Normal}(0, 1) \nonumber\\
\epsilon_i &\sim \mathrm{Normal}(0, 10) \nonumber\\
\beta_0 &= 20 \nonumber\\
\beta_{\mathrm{A}_1} &= 0.1 \nonumber\\
\beta_{\mathrm{A}_2} &= 0.8 \nonumber\\
\beta_{\mathrm{A}_3} &= 0.3 \nonumber\\
\beta_{\mathrm{W}} &= 0.8 \nonumber\\
i &= 1,\cdots, 100
\end{align}
%
We may be interested in comparing the observed average household size ($\over{\mathrm{hh~size}}$) and the predicted average household size ($\over{\widehat{\mathrm{hh~size}}}$). 
%
\begin{align}\label{sim:lm_cubic_mean}
\over{\mathrm{hh~size}} &\approx \beta_0 + \beta_{\mathrm{A_1}}\over{\mathrm{Age}_i} + \beta_{\mathrm{A_2}}\over{\mathrm{Age}^2_i} + \beta_{\mathrm{A_3}}\over{\mathrm{Age}^3_i} \nonumber\\ 
	& + \beta_{\mathrm{W}}\over{\mathrm{Wealthindex}_i}
\end{align}
%

We provide analytical comparison in two cases, i.e., when the interaction is in the focal predictor ($\mathrm{Age}$ is the focal predictor)
%
\begin{align}\label{sim:lm_cubic_analytic_age}
\eta^\star(\mathrm{Age}_i, \nset{{\over{\mathrm{Wealthindex}}}}) &= \beta_0 + \beta_{\mathrm{A_1}}\mathrm{Age}_i + \beta_{\mathrm{A_2}}\mathrm{Age}^2_i + \beta_{\mathrm{A_3}}\mathrm{Age}^3_i \nonumber\\ 
	& + \beta_{\mathrm{W}}\over{\mathrm{Wealthindex}} \nonumber\\
\over{{\eta^\star(\mathrm{Age}_i, \nset{{\over{\mathrm{Wealthindex}}}})}} &= \beta_0 + \beta_{\mathrm{A_1}}\over{{\mathrm{Age}_i}} + \beta_{\mathrm{A_2}}\over{{\mathrm{Age}^2_i}} + \beta_{\mathrm{A_3}}\over{{\mathrm{Age}^3_i}} \nonumber\\ 
	& + \beta_{\mathrm{W}}\over{\mathrm{Wealthindex}}
\end{align}
%
and when the interaction is in the non-focal predictor ($\mathrm{Wealthindex}$ is the focal predictor)
%
\begin{align}\label{sim:lm_cubic_analytic_wealthindex}
\eta^\star(\mathrm{Wealthindex}_i, \nset{{\{\over{\mathrm{Age}}, \over{\mathrm{Age}^2}, \over{\mathrm{Age}^2}\}}}) &= \beta_0 + \beta_{\mathrm{A_1}}\over{\mathrm{Age}} + \beta_{\mathrm{A_2}}\over{\mathrm{Age}^2} + \beta_{\mathrm{A_3}}\over{\mathrm{Age}^3} \nonumber\\ 
	& + \beta_{\mathrm{W}}\mathrm{Wealthindex}_i \nonumber\\
\yE(\eta^\star(\mathrm{Wealthindex}_i, \nset{{\{\over{\mathrm{Age}}, \over{\mathrm{Age}^2}, \over{\mathrm{Age}^3}\}}})) &= \beta_0 + \beta_{\mathrm{A_1}}\over{\mathrm{Age}} + \beta_{\mathrm{A_2}}\over{\mathrm{Age}^2} + \beta_{\mathrm{A_3}}\over{\mathrm{Age}^3} \nonumber\\ 
	& + \beta_{\mathrm{W}}\over{\mathrm{Wealthindex}_i}
\end{align}
%
Depending on the choice of the values of the focal predictors in each case, Equation~\ref{sim:lm_cubic_analytic_age} and \ref{sim:lm_cubic_analytic_wealthindex} approximates \ref{sim:lm_cubic_mean} i.e., mean household size at the model center. In Equation~\ref{sim:lm_cubic_analytic_age}, the model center is not a point in the predictor space but rather described by the terms of the cubic polynomial, as opposed to \ref{sim:lm_cubic_analytic_wealthindex} which has the model center as point.
